Visualization and analysis of probability distributions of large temporal data

### 2021-11-10 Rmd/06-appendix.Rmd
adding chapter 4 supplementary material

```{r external-hakear-supplementary, include = FALSE, cache = FALSE}
read_chunk("scripts/hakear.R")
read_chunk("scripts/hakear-supplementary.R")
knitr::read_chunk('scripts/gracsr.R')
knitr::read_chunk('scripts/gracsr-supplementary.R')
read_chunk(here::here("validation/simdata01.R"))
```

@@

# Chapter 3 supplementary

@@

# Chapter 4 supplementary

```{r raw-data-50, out.width = "100%", fig.cap=" The raw half-hourly energy usage for $50$ sampled households is plotted along the y-axis versus time in a linear scale.  Each of these series is associated with a single customer. It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed. No insightful behavioral pattern could be discerned from this view other than when the customer is not at home."}

knitr::include_graphics("figs/raw_plot_cust.png") # look at smart-meter.R for the code
```

@@

```{r missing-data, fig.cap="(ref:missing-data)"}
knitr::include_graphics("figs/missing-data-5050.png")
```

_Prototype selection_

@@

S4. Run t-SNE using the default arguments on the complete data (sets the perplexity to equal 30 and performs random initialisation). We then create a linked tour with t-SNE layout with R package liminal.

@@

```{r mds, out.width="100%", fig.cap="One instance of brushing in tours (right) and projecting the points in a lower dimensional tsne cloud (left).", out.height="200px"}
knitr::include_graphics("figs/tour-tsne.png")
```

@@

# Interaction of granularities

Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$. In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories. Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs. Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively. D4 varies across both $g1$ and $g2$ categories. D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.

```{r interaction-gran, fig.cap = "The distribution of simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs to extend the proposed validation designs when two granularities of interest interact. D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively. D4 changes across categories of both $g1$ and $g2$."}

@@

```{r JS-mydata-few}
data_q_wide <- data_q %>%
  select(id_facet, id_x, unique_data, sim_data_quantile) %>%
  pivot_wider(names_from = unique_data,
              values_from = sim_data_quantile) %>%
  select(-c(1, 2))

ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
ldata <- nrow(ndata)
lcomb <-  nx_val*nfacet_val

dist_data <- map(1:ldata, function(x){ # first data
  map(1:ldata, function(y){ # 2nd data
    map(1:lcomb, function(z){ # number of combinations nx*nfacet
      JS(
        prob = quantile_prob_val,
        unlist(data_q_wide[z,x]),
        unlist(data_q_wide[z,y])
      ) %>% as_tibble()
    })%>% bind_rows(.id = "combinations")
  })%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

dist_mat <- dist_data %>%
  group_by(data_type1, data_type2) %>%
  summarise(dist = sum(value)) %>%
  pivot_wider(names_from = data_type2,
              values_from = dist) %>%
  mutate(data_type1 = as.numeric(data_type1)) %>%
  left_join(ndata, by = c("data_type1" = "index"))

dist_mat_format <- dist_mat %>%
  ungroup() %>%
  select(-data_type1, -unique_data)

rownames(dist_mat_format) <- dist_mat$unique_data
# write_rds(dist_mat_format, "validation/interact-4D/distmat_40series.rds")
# write_rds(dist_mat, "validation/interact-4D/distmat_40series_withnames.rds")
```

@@

@@

mds_plot_10 <- ggplot(all_data_cluster,
       aes(x = Dim.1,
           y = Dim.2,
       color = groups)) +
  geom_text(label = rownames(mds),check_overlap = TRUE)+
  geom_point(size = 1) +
  theme_classic()+
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "Dark2") +
  coord_fixed(ratio = 1)+
  theme(text = element_text(size = 6))

@@

```{r JS-mydata}
data_q_wide <- data_q %>%
  select(id_facet, id_x, unique_data, sim_data_quantile) %>%
  pivot_wider(names_from = unique_data,
              values_from = sim_data_quantile) %>%
  select(-c(1, 2))

ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
ldata <- nrow(ndata)
lcomb <-  nx_val*nfacet_val

dist_data <- map(1:ldata, function(x){ # first data
  map(1:ldata, function(y){ # 2nd data
    map(1:lcomb, function(z){ # number of combinations nx*nfacet
      JS(
        prob = quantile_prob_val,
        unlist(data_q_wide[z,x]),
        unlist(data_q_wide[z,y])
      ) %>% as_tibble()
    })%>% bind_rows(.id = "combinations")
  })%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

dist_mat <- dist_data %>%
  group_by(data_type1, data_type2) %>%
  summarise(dist = sum(value)) %>%
  pivot_wider(names_from = data_type2,
              values_from = dist) %>%
  mutate(data_type1 = as.numeric(data_type1)) %>%
  left_join(ndata, by = c("data_type1" = "index"))

dist_mat_format <- dist_mat %>%
  ungroup() %>%
  select(-data_type1, -unique_data)

rownames(dist_mat_format) <- dist_mat$unique_data
# write_rds(dist_mat_format, "validation/interact-4D/distmat_400series.rds")
# write_rds(dist_mat, "validation/interact-4D/distmat_400series_withnames.rds")
```

@@

_Clustering all 350 customers_

@@

wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_nqt.rds")) %>% broom::tidy()

moy <- read_rds(here("data/dist_gran_moy_356cust_nqt.rds")) %>% broom::tidy()

hod <- read_rds(here("data/dist_gran_hod_356cust_nqt.rds")) %>% broom::tidy()

@@

data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
  mutate(customer_id = as.character(customer_id)) %>%
  gracsr::scale_gran( method = "robust",
                      response = "general_supply_kwh")

@@

```{r combined-groups-js-entire, fig.cap = "The distribution of electricity demand for the clusters across hod, moy and wkndwday for the $17$ groups from $353$ customers. Wknd-wday variations across groups are not distinguishable, but ideally each group should have an unique combination of hod and moy.", fig.height=6}

>>>>

```{r external-hakear-supplementary, include = FALSE, cache = FALSE}
read_chunk("scripts/hakear.R")
read_chunk("scripts/hakear-supplementary.R")
knitr::read_chunk('scripts/gracsr.R')
knitr::read_chunk('scripts/gracsr-supplementary.R')
```

@@

# Supplementary: Chapter 3

@@

# Supplementary: Chapter 4

## Raw data plot

```{r raw-data-50, out.width = "100%", fig.cap=" The raw half-hourly energy usage for $50$ sampled households is plotted along the y-axis versus time in a linear scale.  Each of these series is associated with a single customer. It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed. No insightful behavioral pattern could be discerned from this view other than when the customer is not at home.", fig.pos="h"}

knitr::include_graphics("img/raw_plot_cust.png") # look at smart-meter.R for the code
```

## Missing data plot

@@

```{r missing-data, fig.cap="(ref:missing-data)", fig.pos="h"}
knitr::include_graphics("img/missing-data-5050.png")
```

## Prototype selection

@@

S4. Run t-SNE using the default arguments on the complete data (sets the perplexity to equal 30 and performs random initialization). We then create a linked tour with t-SNE layout with R package liminal.

@@

```{r mds, out.width="100%", fig.cap="One instance of brushing in tours (right) and projecting the points in a lower dimensional tsne cloud (left).", out.height="200px"}
knitr::include_graphics("img/tour-tsne.png")
```

@@

## Interaction of granularities

Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$. In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories. Consider the $4$ designs in Figure \ref{fig:interaction-gran-supplement}, where various distributions are assumed for different combinations of categories, resulting in different designs. Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively. D4 varies across both $g1$ and $g2$ categories. D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.

```{r interaction-gran-supplement, fig.cap = "The distribution of simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs to extend the proposed validation designs when two granularities of interest interact. D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively. D4 changes across categories of both $g1$ and $g2$."}

@@

```{r JS-mydata-few}
# data_q_wide <- data_q %>%
#   select(id_facet, id_x, unique_data, sim_data_quantile) %>%
#   pivot_wider(names_from = unique_data,
#               values_from = sim_data_quantile) %>%
#   select(-c(1, 2))
# 
# ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
# ldata <- nrow(ndata)
# lcomb <-  nx_val*nfacet_val
# 
# dist_data <- map(1:ldata, function(x){ # first data
#   map(1:ldata, function(y){ # 2nd data
#     map(1:lcomb, function(z){ # number of combinations nx*nfacet
#       JS(
#         prob = quantile_prob_val,
#         unlist(data_q_wide[z,x]),
#         unlist(data_q_wide[z,y])
#       ) %>% as_tibble()
#     })%>% bind_rows(.id = "combinations")
#   })%>% bind_rows(.id = "data_type1")
# }) %>% bind_rows(.id = "data_type2") 
# 
# 
# 
# dist_mat <- dist_data %>%
#   group_by(data_type1, data_type2) %>%
#   summarise(dist = sum(value)) %>%
#   pivot_wider(names_from = data_type2,
#               values_from = dist) %>%
#   mutate(data_type1 = as.numeric(data_type1)) %>%
#   left_join(ndata, by = c("data_type1" = "index"))
# 
# 
# dist_mat_format <- dist_mat %>%
#   ungroup() %>%
#   select(-data_type1, -unique_data)
# 
# 
# rownames(dist_mat_format) <- dist_mat$unique_data
# write_rds(dist_mat_format, "data/gracsr/validation/interact-4D/distmat_40series.rds")
# write_rds(dist_mat, "data/gracsr/validation/interact-4D/distmat_40series_withnames.rds")
# 
dist_mat_format <- read_rds("data/gracsr/validation/interact-4D/distmat_40series.rds")
dist_mat <- read_rds("data/gracsr/validation/interact-4D/distmat_40series_withnames.rds")
```

@@

colnames(mds) <- c("Dim.1", "Dim.2")
#rownames(mds) <- dist_mat$unique_data

@@

mds_plot_10 <- ggplot(all_data_cluster,
       aes(x = Dim.1,
           y = Dim.2,
       color = groups)) +
  geom_text(label = rownames(dist_mat_format),check_overlap = TRUE)+
  geom_point(size = 1) +
  theme_classic()+
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "Dark2") +
  coord_fixed(ratio = 1)+
  theme(text = element_text(size = 6))

@@

```{r JS-mydata}
# data_q_wide <- data_q %>%
#   select(id_facet, id_x, unique_data, sim_data_quantile) %>%
#   pivot_wider(names_from = unique_data,
#               values_from = sim_data_quantile) %>%
#   select(-c(1, 2))
# 
# ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
# ldata <- nrow(ndata)
# lcomb <-  nx_val*nfacet_val
# 
# dist_data <- map(1:ldata, function(x){ # first data
#   map(1:ldata, function(y){ # 2nd data
#     map(1:lcomb, function(z){ # number of combinations nx*nfacet
#       JS(
#         prob = quantile_prob_val,
#         unlist(data_q_wide[z,x]),
#         unlist(data_q_wide[z,y])
#       ) %>% as_tibble()
#     })%>% bind_rows(.id = "combinations")
#   })%>% bind_rows(.id = "data_type1")
# }) %>% bind_rows(.id = "data_type2")
# 
# 
# 
# dist_mat <- dist_data %>%
#   group_by(data_type1, data_type2) %>%
#   summarise(dist = sum(value)) %>%
#   pivot_wider(names_from = data_type2,
#               values_from = dist) %>%
#   mutate(data_type1 = as.numeric(data_type1)) %>%
#   left_join(ndata, by = c("data_type1" = "index"))
# 
# 
# dist_mat_format <- dist_mat %>%
#   ungroup() %>%
#   select(-data_type1, -unique_data)
# 
# 
# rownames(dist_mat_format) <- dist_mat$unique_data
# write_rds(dist_mat_format, "data/gracsr/validation/interact-4D/distmat_400series.rds")
# write_rds(dist_mat, "data/gracsr/validation/interact-4D/distmat_400series_withnames.rds")
dist_mat_format <- read_rds("data/gracsr/validation/interact-4D/distmat_400series.rds")
dist_mat <- read_rds("data/gracsr/validation/interact-4D/distmat_400series_withnames.rds")

@@

## Clustering 353 customers

<!-- All the customers after data filtering and variable selection in the main chapter, before proceeding to instance selection. -->

@@

wkndwday <- read_rds(here("data/gracsr/dist_gran_wkndwday_356cust_nqt.rds")) %>% broom::tidy()

moy <- read_rds(here("data/gracsr/dist_gran_moy_356cust_nqt.rds")) %>% broom::tidy()

hod <- read_rds(here("data/gracsr/dist_gran_hod_356cust_nqt.rds")) %>% broom::tidy()

@@

data_pick <- read_rds(here::here("data/gracsr/elec_nogap_2013_clean_356cust.rds")) %>%
  mutate(customer_id = as.character(customer_id)) %>%
  gracsr::scale_gran( method = "robust",
                      response = "general_supply_kwh")

@@

```{r combined-groups-js-entire-all, fig.cap = "The distribution of electricity demand for the clusters across hod, moy and wkndwday for the $17$ groups from $353$ customers. Wknd-wday variations across groups are not distinguishable, but ideally each group should have an unique combination of hod and moy.", fig.height=6}
